Type Refinements for the Working Class

### 2015-11-22 paper.tex
Comment out the bulk of the paper pending theory modifications

We will require two judgments, defined mutually: $\RefinesCtx{\Phi}{\Upsilon}$,
which means that $\Phi$ is a symbol context which refines the symbol context
$\Upsilon$ (presupposing $\IsSymCtx{\Upsilon}$), and parametric refinement
$\Refines{\Phi}{\phi}{\tau}$ (which shall presuppose
$\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$).

\newcommand\RefinesCtxNil{\sqsubset^\star_{\mathtt{nil}}}
\newcommand\RefinesCtxSnoc{\sqsubset^\star_{\mathtt{snoc}}}

The first judgment we define inductively:
\begin{gather*}
  \infer[\RefinesCtxNil]{
    \RefinesCtx{\cdot}{\cdot}
  }{
  }\qquad
  \infer[\RefinesCtxSnoc]{
    \RefinesCtx{\Phi,u:\phi}{\Upsilon,u:\tau}
  }{
    \RefinesCtx{\Phi}{\Upsilon} &
    \Refines{\Phi}{\phi}{\tau}
  }
\end{gather*}

The second judgment is defined coinductively via the following meaning
explanation:
\begin{definition}[Parametric Refinement]
  To know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
  $\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$) is to know, for any renaming
  $\Of{\varrho}{\Upsilon\inj\OMode{\Upsilon'}}$, and for any $M, N$,
  what it means for $M$ and $N$ to be identified under $\phi$, supposing
  $\IsAbt{\cdot}{\Upsilon'}{\cdot}{M}{\tau}$ and
  $\IsAbt{\cdot}{\Upsilon'}{\cdot}{N}{\tau}$, requiring that if
  $\IMode{M}\sim^\rho_\phi\IMode{N}$ then $\IMode{N}\sim^\rho_\phi\IMode{M}$, and if
  $\IMode{M}\sim^\rho_\phi\IMode{N}$ and $\IMode{N}\sim^\rho_\phi\IMode{O}$, then
  $\IMode{M}\sim^\rho_\phi\IMode{O}$.
\end{definition}

In other words, to know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
$\RefinesCtx{\Phi}{\Upsilon}$) is to know, for any renaming of $\Upsilon$ to
$\Upsilon'$, which partial equivalence relation on the closed $\tau$-sorted
$\Upsilon'$-terms $\phi$ denotes. At this point, it should be remarked that
this is very similar to the semantical explanation of typehood given in
\cite{martin-lof:1979}, except that we have generalized it to a multi-sorted
setting, and that we have fibred the entire apparatus over collections
$\Upsilon$ of symbols.

\todo[inline]{%
  define maps of refined contexts, then change weakening lemma to
  monotonicity, which is stronger and more useful.
}
\begin{thm}[Weakening]
  If $\Refines{\Phi}{\phi}{\tau}$, then $\Refines{\Phi,u:\psi}{\phi}{\tau}$ (supposing
  $\NotIn{u}{\Dom\Phi}$ and $\Refines{\Phi}{\psi}{\sigma}$).
\end{thm}
\begin{proof}
  This follows trivially from the rule $\RefinesCtxSnoc$ and the canonical renaming
  $\Upsilon\inj\Upsilon,u:\sigma$.
\end{proof}

When we have $\Refines{\cdot}{\phi}{\tau}$, we say that $\phi$ \emph{globally}
refines $\tau$ and may simply write $\GRefines{\phi}{\tau}$. By weakening, if
$\GRefines{\phi}{\tau}$ then for all $\Phi$, we have
$\Refines{\Phi}{\phi}{\tau}$ supposing $\RefinesCtx{\Phi}{\Upsilon}$.

\subsubsection{Ordering and Equality of Parametric Refinement}

We will write $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ to mean that $\phi$ is a
\emph{subrefinement} of $\psi$, presupposing $\RefinesCtx{\Phi}{\Upsilon}$,
$\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$:
\begin{definition}
  To know $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ is to know, for any substitution
  $\Of{\rho}{\Upsilon\inj\OMode{\Upsilon'}}$, the following judgments:
  \[
    \GenJ{
      \MV{m},\MV{n}:\tau
    }{
      \HypJ{
        \IMode{M}\sim^\rho_\psi\IMode{N}
      }{
        \IMode{M}\sim^\rho_\phi\IMode{N}
      }
    }
  \]
\end{definition}

Two refinements are equal when they denote the same PER. That is, we have
$\EqRefines{\Phi}{\phi}{\psi}{\tau}$ just when both
$\SubRefines{\Phi}{\phi}{\psi}{\tau}$ and $\SubRefines{\Phi}{\psi}{\phi}{\tau}$.

\subsection{Parametric Equality}

The primary judgment concerning refinements is the parametric equality,
$\RMemEq{\Phi}{M}{N}{\phi}$, which presupposes $\RefinesCtx{\Phi}{\Upsilon}$,
$\Refines{\Phi}{\phi}{\tau}$, $\IsAbt{\cdot}{\Upsilon}{\cdot}{M}{\tau}$ and
$\IsAbt{\cdot}{\Upsilon}{\cdot}{N}{\tau}$. The meaning of this judgment is that
$M$ and $N$ are identified by $\phi$ at the identity renaming
$\Of{1}{\IMode{\Upsilon}\inj\OMode{\Upsilon}}$:
\[
  \infer{
    \RMemEq{\Phi}{M}{N}{\phi}
  }{
    \IMode{M}\sim^1_\phi\IMode{N}
  }
\]

\subsection{Functional Refinement}

\newcommand\PRefinesCtxNil{\leftidx{^\parallel}{\!\!\sqsubset}{^\star_{\mathtt{nil}}}}
\newcommand\PRefinesCtxSnoc{\leftidx{^\parallel}{\!\!\sqsubset}{^\star_{\mathtt{snoc}}}}

Next, we define functional refinement
$\HEqRefines{\Phi}{\Psi}{\phi}{\psi}{\tau}$ in terms of parametric refinement,
simultaneously with parametric context refinement
$\PRefinesCtx{\Phi}{\Psi}{\Gamma}$ and functional equality
$\HRMemEq{\Phi}{\Psi}{M}{N}{\phi}$. We will write
$\HRefines{\Phi}{\Psi}{\phi}{\tau}$ as a shorthand for
$\HEqRefines{\Phi}{\Psi}{\phi}{\phi}{\tau}$.

\newcommand\DefPRefinesCtxNil[1]{
  \infer[\PRefinesCtxNil]{
    \PRefinesCtx{#1}{\cdot}{\cdot}
  }{
  }
}
\newcommand\DefPRefinesCtxSnoc[6]{
  \infer[\PRefinesCtxSnoc]{
    \PRefinesCtx{#1}{#2,#3:#4}{#5,#3:#6}
  }{
    \PRefinesCtx{#1}{#2}{#5} &
    \HRefines{#1}{#2}{#4}{#6}
  }
}

Parametric context refinement $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$ shall be
defined as follows, presupposing $\RefinesCtx{\Phi}{\Upsilon}$ and
$\IsVarCtx{\Gamma}$:
\begin{gather*}
  \DefPRefinesCtxNil{\Phi}\qquad
  \DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\phi}{\Gamma}{\tau}
\end{gather*}

Functional refinement $\HEqRefines{\Phi}{\Psi}{\phi}{\psi}{\tau}$ presupposes
$\PRefinesCtx{\Phi}{\Psi}{\Gamma}$ and $\IsSort{\tau}$, and is explained by
induction on the evidence for its first presupposition.

\paragraph{Case} $\DefPRefinesCtxNil{\Phi}$.
\[
  \infer{
    \HEqRefines{\Phi}{\cdot}{\phi}{\psi}{\tau}
  }{
    \EqRefines{\Phi}{\phi}{\psi}{\tau}
  }
\]

\paragraph{Case} $\DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\chi}{\Gamma}{\sigma}$,
where $\RefinesCtx{\Phi}{\Upsilon}$.

\begin{gather*}
  \infer{
    \HEqRefines{\Phi}{\Psi,x:\chi}{\phi}{\psi}{\tau}
  }{
    \HPGenJ{\Upsilon}{\Gamma}{\MV{m},\MV{n}:\sigma}{
      \HypJ{
        \HEqRefines{\Phi}{\Psi}{
          \Subst{\MV{m}}{x}{\phi}
        }{
          \Subst{\MV{n}}{x}{\psi}
        }{\tau}
      }{
        \HRMemEq{\Phi}{\Psi}{\MV{n}}{\MV{n}}{\chi}
      }
    }
  }
\end{gather*}

\subsection{Functional Equality}

Functional equality $\HRMemEq{\Phi}{\Psi}{M}{N}{\phi}$ presupposes
$\RefinesCtx{\Phi}{\Upsilon}$, $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$,
$\HRefines{\Phi}{\Psi}{\phi}{\tau}$, $\IsAbt{\cdot}{\Upsilon}{\Gamma}{M}{\tau}$
and $\IsAbt{\cdot}{\Upsilon}{\Gamma}{N}{\tau}$; the judgment is defined by
induction on the evidence for the second presupposition
$\PRefinesCtx{\Phi}{\Psi}{\Gamma}$:

\paragraph{Case} $\DefPRefinesCtxNil{\Phi}$.
\[
  \infer{
    \HRMemEq{\Phi}{\cdot}{M}{N}{\phi}
  }{
    \RMemEq{\Phi}{M}{N}{\phi}
  }
\]

\paragraph{Case} $\DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\psi}{\Gamma}{\sigma}$,
where $\RefinesCtx{\Phi}{\Upsilon}$.
\[
  \infer{
    \HRMemEq{\Phi}{\Psi,x:\psi}{M}{N}{\phi}
  }{
    \HPGenJ{\Upsilon}{\Gamma}{\MV{m},\MV{n}:\sigma}{
      \HypJ{
        \HRMemEq{\Phi}{\Psi}{
          \Subst{\MV{m}}{x}{\phi}
        }{
          \Subst{\MV{n}}{x}{\psi}
        }{\phi}
      }{
        \HRMemEq{\Phi}{\Psi}{\MV{m}}{\MV{n}}{\psi}
      }
    }
  }
\]

\section{Constructions on Refinements}

Refinements can be arranged in a partial order via the
$\SubRefines{\Phi}{\phi}{\psi}{\tau}$ judgment, and admit a lattice structure
at any sort $\tau$.

We have the bottom refinement $\GRefines{\TyBot{\tau}}{\tau}$, which simply
contains no elements. Next, the top refinement $\GRefines{\TyTop{\tau}}{\tau}$
identifies all terms $M,N$ such that $\IsAbt{\cdot}{\Upsilon}{\cdot}{M}{\tau}$
and $\IsAbt{\cdot}{\Upsilon}{\cdot}{N}{\tau}$. The join
$\Refines{\Phi}{\phi\TyJoin{\tau}\psi}{\tau}$ of two refinements
$\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$ is defined as follows:
\begin{gather*}
  \infer{
    \RMemEq{\Phi}{M}{N}{\phi\TyJoin{\tau}\psi}
  }{
    \RMemEq{\Phi}{M}{N}{\phi}
  }\qquad
  \infer{
    \RMemEq{\Phi}{M}{N}{\phi\TyJoin{\tau}\psi}
  }{
    \RMemEq{\Phi}{M}{N}{\psi}
  }
\end{gather*}

Finally the meet $\Refines{\Phi}{\phi\TyMeet{\tau}\psi}{\tau}$ of two refinements
$\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$ is defined
dually:
\[
  \infer{
    \RMemEq{\Phi}{M}{N}{\phi\TyMeet{\tau}\psi}
  }{
    \RMemEq{\Phi}{M}{N}{\phi} &
    \RMemEq{\Phi}{M}{N}{\psi}
  }
\]

It is trivial to verify that these constructions on refinements give rise to a
lattice.

\section{Multi-Sorted Computational Type Theory}

Martin-L\"of's type theory is a theory of \emph{value types}: to define a type,
you specify how to form its equal canonical inhabitants, or values. Then, the
inhabitants of the types are explained uniformly by extending this relation
over an operational semantics such that two terms are equal members of a
type just when they compute to equal values of that type.

More generally, one may wish to consider types which include computations in
addition to values; an example of such a type would be $\bar{\psi}$ when $\psi$
is a type, which would include all the members of $\psi$ as well as a divergent
computation. Another example is the type $\TyTop{\tau}$ which identifies all
closed terms of sort $\tau$.

In order to ensure that such types are sensible, we must formulate a notion of
computational equivalence, and then say that a refinement is a \emph{type} just
when it respects computational equivalence; then, we are able to define types
as PERs on \emph{programs} in addition to values.

>>>>

\todo[inline]{%
  Restore definitions of the refinement framework after they have been fixed!!
}

%% \subsection{Parametric Refinement}
%%
%% We will require two judgments, defined mutually: $\RefinesCtx{\Phi}{\Upsilon}$,
%% which means that $\Phi$ is a symbol context which refines the symbol context
%% $\Upsilon$ (presupposing $\IsSymCtx{\Upsilon}$), and parametric refinement
%% $\Refines{\Phi}{\phi}{\tau}$ (which shall presuppose
%% $\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$).
%%
%% \newcommand\RefinesCtxNil{\sqsubset^\star_{\mathtt{nil}}}
%% \newcommand\RefinesCtxSnoc{\sqsubset^\star_{\mathtt{snoc}}}
%%
%% The first judgment we define inductively:
%% \begin{gather*}
%%   \infer[\RefinesCtxNil]{
%%     \RefinesCtx{\cdot}{\cdot}
%%   }{
%%   }\qquad
%%   \infer[\RefinesCtxSnoc]{
%%     \RefinesCtx{\Phi,u:\phi}{\Upsilon,u:\tau}
%%   }{
%%     \RefinesCtx{\Phi}{\Upsilon} &
%%     \Refines{\Phi}{\phi}{\tau}
%%   }
%% \end{gather*}
%%
%% The second judgment is defined coinductively via the following meaning
%% explanation:
%% \begin{definition}[Parametric Refinement]
%%   To know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
%%   $\RefinesCtx{\Phi}{\Upsilon}$ and $\IsSort{\tau}$) is to know, for any renaming
%%   $\Of{\varrho}{\Upsilon\inj\OMode{\Upsilon'}}$, and for any $M, N$,
%%   what it means for $M$ and $N$ to be identified under $\phi$, supposing
%%   $\IsAbt{\cdot}{\Upsilon'}{\cdot}{M}{\tau}$ and
%%   $\IsAbt{\cdot}{\Upsilon'}{\cdot}{N}{\tau}$, requiring that if
%%   $\IMode{M}\sim^\rho_\phi\IMode{N}$ then $\IMode{N}\sim^\rho_\phi\IMode{M}$, and if
%%   $\IMode{M}\sim^\rho_\phi\IMode{N}$ and $\IMode{N}\sim^\rho_\phi\IMode{O}$, then
%%   $\IMode{M}\sim^\rho_\phi\IMode{O}$.
%% \end{definition}
%%
%% In other words, to know $\Refines{\Phi}{\phi}{\tau}$ (presupposing
%% $\RefinesCtx{\Phi}{\Upsilon}$) is to know, for any renaming of $\Upsilon$ to
%% $\Upsilon'$, which partial equivalence relation on the closed $\tau$-sorted
%% $\Upsilon'$-terms $\phi$ denotes. At this point, it should be remarked that
%% this is very similar to the semantical explanation of typehood given in
%% \cite{martin-lof:1979}, except that we have generalized it to a multi-sorted
%% setting, and that we have fibred the entire apparatus over collections
%% $\Upsilon$ of symbols.
%%
%% \todo[inline]{%
%%   define maps of refined contexts, then change weakening lemma to
%%   monotonicity, which is stronger and more useful.
%% }
%% \begin{thm}[Weakening]
%%   If $\Refines{\Phi}{\phi}{\tau}$, then $\Refines{\Phi,u:\psi}{\phi}{\tau}$ (supposing
%%   $\NotIn{u}{\Dom\Phi}$ and $\Refines{\Phi}{\psi}{\sigma}$).
%% \end{thm}
%% \begin{proof}
%%   This follows trivially from the rule $\RefinesCtxSnoc$ and the canonical renaming
%%   $\Upsilon\inj\Upsilon,u:\sigma$.
%% \end{proof}
%%
%% When we have $\Refines{\cdot}{\phi}{\tau}$, we say that $\phi$ \emph{globally}
%% refines $\tau$ and may simply write $\GRefines{\phi}{\tau}$. By weakening, if
%% $\GRefines{\phi}{\tau}$ then for all $\Phi$, we have
%% $\Refines{\Phi}{\phi}{\tau}$ supposing $\RefinesCtx{\Phi}{\Upsilon}$.
%%
%% \subsubsection{Ordering and Equality of Parametric Refinement}
%%
%% We will write $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ to mean that $\phi$ is a
%% \emph{subrefinement} of $\psi$, presupposing $\RefinesCtx{\Phi}{\Upsilon}$,
%% $\Refines{\Upsilon}{\phi}{\tau}$ and $\Refines{\Upsilon}{\psi}{\tau}$:
%% \begin{definition}
%%   To know $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ is to know, for any substitution
%%   $\Of{\rho}{\Upsilon\inj\OMode{\Upsilon'}}$, the following judgments:
%%   \[
%%     \GenJ{
%%       \MV{m},\MV{n}:\tau
%%     }{
%%       \HypJ{
%%         \IMode{M}\sim^\rho_\psi\IMode{N}
%%       }{
%%         \IMode{M}\sim^\rho_\phi\IMode{N}
%%       }
%%     }
%%   \]
%% \end{definition}
%%
%% Two refinements are equal when they denote the same PER. That is, we have
%% $\EqRefines{\Phi}{\phi}{\psi}{\tau}$ just when both
%% $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ and $\SubRefines{\Phi}{\psi}{\phi}{\tau}$.
%%
%% \subsection{Parametric Equality}
%%
%% The primary judgment concerning refinements is the parametric equality,
%% $\RMemEq{\Phi}{M}{N}{\phi}$, which presupposes $\RefinesCtx{\Phi}{\Upsilon}$,
%% $\Refines{\Phi}{\phi}{\tau}$, $\IsAbt{\cdot}{\Upsilon}{\cdot}{M}{\tau}$ and
%% $\IsAbt{\cdot}{\Upsilon}{\cdot}{N}{\tau}$. The meaning of this judgment is that
%% $M$ and $N$ are identified by $\phi$ at the identity renaming
%% $\Of{1}{\IMode{\Upsilon}\inj\OMode{\Upsilon}}$:
%% \[
%%   \infer{
%%     \RMemEq{\Phi}{M}{N}{\phi}
%%   }{
%%     \IMode{M}\sim^1_\phi\IMode{N}
%%   }
%% \]
%%
%% \subsection{Functional Refinement}
%%
%% \newcommand\PRefinesCtxNil{\leftidx{^\parallel}{\!\!\sqsubset}{^\star_{\mathtt{nil}}}}
%% \newcommand\PRefinesCtxSnoc{\leftidx{^\parallel}{\!\!\sqsubset}{^\star_{\mathtt{snoc}}}}
%%
%% Next, we define functional refinement
%% $\HEqRefines{\Phi}{\Psi}{\phi}{\psi}{\tau}$ in terms of parametric refinement,
%% simultaneously with parametric context refinement
%% $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$ and functional equality
%% $\HRMemEq{\Phi}{\Psi}{M}{N}{\phi}$. We will write
%% $\HRefines{\Phi}{\Psi}{\phi}{\tau}$ as a shorthand for
%% $\HEqRefines{\Phi}{\Psi}{\phi}{\phi}{\tau}$.
%%
%% \newcommand\DefPRefinesCtxNil[1]{
%%   \infer[\PRefinesCtxNil]{
%%     \PRefinesCtx{#1}{\cdot}{\cdot}
%%   }{
%%   }
%% }
%% \newcommand\DefPRefinesCtxSnoc[6]{
%%   \infer[\PRefinesCtxSnoc]{
%%     \PRefinesCtx{#1}{#2,#3:#4}{#5,#3:#6}
%%   }{
%%     \PRefinesCtx{#1}{#2}{#5} &
%%     \HRefines{#1}{#2}{#4}{#6}
%%   }
%% }
%%
%% Parametric context refinement $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$ shall be
%% defined as follows, presupposing $\RefinesCtx{\Phi}{\Upsilon}$ and
%% $\IsVarCtx{\Gamma}$:
%% \begin{gather*}
%%   \DefPRefinesCtxNil{\Phi}\qquad
%%   \DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\phi}{\Gamma}{\tau}
%% \end{gather*}
%%
%% Functional refinement $\HEqRefines{\Phi}{\Psi}{\phi}{\psi}{\tau}$ presupposes
%% $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$ and $\IsSort{\tau}$, and is explained by
%% induction on the evidence for its first presupposition.
%%
%% \paragraph{Case} $\DefPRefinesCtxNil{\Phi}$.
%% \[
%%   \infer{
%%     \HEqRefines{\Phi}{\cdot}{\phi}{\psi}{\tau}
%%   }{
%%     \EqRefines{\Phi}{\phi}{\psi}{\tau}
%%   }
%% \]
%%
%% \paragraph{Case} $\DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\chi}{\Gamma}{\sigma}$,
%% where $\RefinesCtx{\Phi}{\Upsilon}$.
%%
%% \begin{gather*}
%%   \infer{
%%     \HEqRefines{\Phi}{\Psi,x:\chi}{\phi}{\psi}{\tau}
%%   }{
%%     \HPGenJ{\Upsilon}{\Gamma}{\MV{m},\MV{n}:\sigma}{
%%       \HypJ{
%%         \HEqRefines{\Phi}{\Psi}{
%%           \Subst{\MV{m}}{x}{\phi}
%%         }{
%%           \Subst{\MV{n}}{x}{\psi}
%%         }{\tau}
%%       }{
%%         \HRMemEq{\Phi}{\Psi}{\MV{n}}{\MV{n}}{\chi}
%%       }
%%     }
%%   }
%% \end{gather*}
%%
%% \subsection{Functional Equality}
%%
%% Functional equality $\HRMemEq{\Phi}{\Psi}{M}{N}{\phi}$ presupposes
%% $\RefinesCtx{\Phi}{\Upsilon}$, $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$,
%% $\HRefines{\Phi}{\Psi}{\phi}{\tau}$, $\IsAbt{\cdot}{\Upsilon}{\Gamma}{M}{\tau}$
%% and $\IsAbt{\cdot}{\Upsilon}{\Gamma}{N}{\tau}$; the judgment is defined by
%% induction on the evidence for the second presupposition
%% $\PRefinesCtx{\Phi}{\Psi}{\Gamma}$:
%%
%% \paragraph{Case} $\DefPRefinesCtxNil{\Phi}$.
%% \[
%%   \infer{
%%     \HRMemEq{\Phi}{\cdot}{M}{N}{\phi}
%%   }{
%%     \RMemEq{\Phi}{M}{N}{\phi}
%%   }
%% \]
%%
%% \paragraph{Case} $\DefPRefinesCtxSnoc{\Phi}{\Psi}{x}{\psi}{\Gamma}{\sigma}$,
%% where $\RefinesCtx{\Phi}{\Upsilon}$.
%% \[
%%   \infer{
%%     \HRMemEq{\Phi}{\Psi,x:\psi}{M}{N}{\phi}
%%   }{
%%     \HPGenJ{\Upsilon}{\Gamma}{\MV{m},\MV{n}:\sigma}{
%%       \HypJ{
%%         \HRMemEq{\Phi}{\Psi}{
%%           \Subst{\MV{m}}{x}{\phi}
%%         }{
%%           \Subst{\MV{n}}{x}{\psi}
%%         }{\phi}
%%       }{
%%         \HRMemEq{\Phi}{\Psi}{\MV{m}}{\MV{n}}{\psi}
%%       }
%%     }
%%   }
%% \]
%%
%% \section{Constructions on Refinements}
%%
%% Refinements can be arranged in a partial order via the
%% $\SubRefines{\Phi}{\phi}{\psi}{\tau}$ judgment, and admit a lattice structure
%% at any sort $\tau$.
%%
%% We have the bottom refinement $\GRefines{\TyBot{\tau}}{\tau}$, which simply
%% contains no elements. Next, the top refinement $\GRefines{\TyTop{\tau}}{\tau}$
%% identifies all terms $M,N$ such that $\IsAbt{\cdot}{\Upsilon}{\cdot}{M}{\tau}$
%% and $\IsAbt{\cdot}{\Upsilon}{\cdot}{N}{\tau}$. The join
%% $\Refines{\Phi}{\phi\TyJoin{\tau}\psi}{\tau}$ of two refinements
%% $\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$ is defined as follows:
%% \begin{gather*}
%%   \infer{
%%     \RMemEq{\Phi}{M}{N}{\phi\TyJoin{\tau}\psi}
%%   }{
%%     \RMemEq{\Phi}{M}{N}{\phi}
%%   }\qquad
%%   \infer{
%%     \RMemEq{\Phi}{M}{N}{\phi\TyJoin{\tau}\psi}
%%   }{
%%     \RMemEq{\Phi}{M}{N}{\psi}
%%   }
%% \end{gather*}
%%
%% Finally the meet $\Refines{\Phi}{\phi\TyMeet{\tau}\psi}{\tau}$ of two refinements
%% $\Refines{\Phi}{\phi}{\tau}$ and $\Refines{\Phi}{\psi}{\tau}$ is defined
%% dually:
%% \[
%%   \infer{
%%     \RMemEq{\Phi}{M}{N}{\phi\TyMeet{\tau}\psi}
%%   }{
%%     \RMemEq{\Phi}{M}{N}{\phi} &
%%     \RMemEq{\Phi}{M}{N}{\psi}
%%   }
%% \]
%%
%% It is trivial to verify that these constructions on refinements give rise to a
%% lattice.
%%
%% \section{Multi-Sorted Computational Type Theory}
%%
%% Martin-L\"of's type theory is a theory of \emph{value types}: to define a type,
%% you specify how to form its equal canonical inhabitants, or values. Then, the
%% inhabitants of the types are explained uniformly by extending this relation
%% over an operational semantics such that two terms are equal members of a
%% type just when they compute to equal values of that type.
%%
%% More generally, one may wish to consider types which include computations in
%% addition to values; an example of such a type would be $\bar{\psi}$ when $\psi$
%% is a type, which would include all the members of $\psi$ as well as a divergent
%% computation. Another example is the type $\TyTop{\tau}$ which identifies all
%% closed terms of sort $\tau$.
%%
%% In order to ensure that such types are sensible, we must formulate a notion of
%% computational equivalence, and then say that a refinement is a \emph{type} just
%% when it respects computational equivalence; then, we are able to define types
%% as PERs on \emph{programs} in addition to values.
